\documentclass[letterpaperr,12pt]{article}

\usepackage[utf8]{inputenc}
\usepackage{amsmath,amssymb}

%\usepackage[spanish]{babel}

\usepackage[T1]{fontenc}  
\usepackage{textcomp}


%\usepackage[garamond,cmbraces]{newtxmath}

\usepackage{xspace}
\usepackage{graphicx}

\usepackage{float}
\usepackage{caption}
\usepackage{subcaption}
%\usepackage{anysize}
\usepackage[margin=2cm]{geometry}

%\usepackage{bm}

\bibliographystyle{alpha}

\newcommand{\Acase}{\textbf{A}\xspace}
\newcommand{\Bcase}{\textbf{B}\xspace}
\newcommand{\Ccase}{\textbf{C}\xspace}
\newcommand{\Dcase}{\textbf{D}\xspace}
\newcommand{\Uno}{\textbf{1}\xspace}
\newcommand{\Dos}{\textbf{2}\xspace}

\renewcommand{\rmdefault}{ugm}
  

\title{Unrolling Fick-Jacobs}
\author{W. P. K. Zapfe}

\abstract{ Due to the apparent impossibility to explain with
  Szabos's Formula the 
 first encounter times in the problem of the two Brownian 
 disks inside a rectangular container, we proceed to adapt the 
 develpment of Vazques and Dagdug for Fick-Jacobs' Formula.}


\section{The Story so far}

The application of Szabo's formula to the problem of
the \emph{contagious disks}  assumed that it was applicable
to consider the avaible free space as essentially equivalent to
the same space distributed inside more simmetrical shapes. 
To be precisse, we were trying to approximate an hypercube with
a fourdimensional absorving rod placed inside it by 
an anular section and two irrelevant ``heights''. We
have trustworthy arguments to support the idea that if this
approximation would have been the real system under study, 
Szabo's, David and Luca Formulas would have applied without
any significant modification. 

To apply Szabo's two dimensional formula, we needed an equivalent
anular geometry. Our approach was to obtain interior and exterior
radii by preserving the aviable free space and absorving surface. 
The extra dimensional ``heights'' of the cilinder turn out to be
unimportant. This approach lead us to an effective interior radious
as function of the disks original radii. Due to the
diagonal placement of this absorving object inside the hypercube,
the formula had to account for pointy tips, which were more
relevant as the radious became larger.  The rather unconfortable
formula accounted correctly for all these geometric conundrums. 

Once in possesion of an effective inner and outer radious in function
of the disks radius ($\rho(r)$) we could apply Szabo's Result. 
It shows a rather systematic discrepance with our numerical results. 
Being that these Numerical Simulations where done by two different
persons using diferent set of dynamic rules with different programing
styles and  languages, and that these coincide rather well, we are
lead to believe that these Numerics represent adecuatelly the 
\emph{mean encounter time} of the two disks, and that our reasoning
for the application of Szabo's is wrong, or is simply not
applicable.

The effective inner radius of the annulus is given in the formula below,
an the figure following shows our simulations (David, Karel) 
against Szabo's Formula
with this expression used. 


\section{ Conjecture about the geometry}.
 
We could make an intermediary approximation to the problem.
In this approximation we still shall use the rotational symetry, that
is, we shall still have sections with anular free space. But we shall preserve
from the original problems the ``narrowing'' of the free space along
the axis of the absorving cilinder. This gives us  
conic anular approximation. This is easy to represent in the 3d case, and
a bit more complicated in the 4d case, but nonetheless is the same
geometry. In this case Szabo's approximation would not be valid as it is, 
due to the narrowing of the free space. In order to test our conjecture
first we present simulations in this geometry and compare it to 
the ``animals in a territory'' simulations. 
If these are aceptably close, we may try to use a Fick-Jacobs approach here.





\begin{document}

\maketitle


\end{document}